---
title: "Data processing"
author: "Yefrid Cordoba"
format: html
---

First import the table with weights and thickness

```{r}
#| message: False
#| warning: False
library(tidyverse)
library(readxl)
library(stringr)
library(janitor)
library(units)
library(emmeans)
library(lmerTest)
library(lme4)
library(performance)
library(see)
library(DHARMa)
library(pbkrtest)
```

Now we start with the data 

```{r}
fines_data <- read_xlsx("raw_data/Fines_study.xlsx", 
                        sheet = "Handsheet_properties")
```

Extract the pulping data, and just select the kappa number

Here i can select which columns are going to appear in my final table
```{r}
Pulping_data <- read_xlsx("raw_data/Fines_study.xlsx", 
                        sheet = "Pulping") #|>
  # select(Biomass, Code, Kappa, Total_screened_yield, )
```

Here i can select which columns are going to appear in my final table
```{r}
freeness_data <- read_xlsx("raw_data/Fines_study.xlsx", 
                        sheet = "Freeness") |>
  select(ID1, freeness)
```

```{r}
Pulping_data
```

Getting the data from the tensile tests

```{r}
tensile_data <- read_xlsx("raw_data/tensile_data.xlsx")
tensile_data
```


We need to create the variables for each of the columns
```{r}
all_data <- fines_data |>
  mutate(biomass = as.factor(Biomass),
         Biomass = NULL,
         fines = factor(ifelse(
           str_detect(ID,"F"), "Y", "N")),
         hs_treatment = factor(
           case_when(
            str_detect(ID, "A") ~ "additive",
            str_detect(ID, "5") ~ "recirculation",
            str_detect(ID, "[1-4]") ~ "control"),
           levels = c("control", "recirculation", "additive")),
         Code = str_extract(ID, "^[WE]\\d"),
         ID1 = str_extract(ID, "^[WE]\\d.")
         ) |>
  inner_join(Pulping_data, by = join_by(Code)) |>
  left_join(freeness_data, by = join_by(ID1)) |>
  left_join(tensile_data, by = join_by(ID)) |>
  select(-Biomass, -ID1) |>
  relocate(Code, biomass, fines, hs_treatment, .after = ID) |>
  mutate(
    id2 = str_extract(str_remove(ID, "A"), "^.{3}")
  )
  
```

Now we get a code to extract the data from the FQA values

```{r}
fqa_data <- readRDS("raw_data/fqa_data.rds")

```

Now that all the data is in that table, we need to fix the data

```{r}
fqa2_data <- fqa_data|>
  slice(-c(1:13, 19, 22, 23, 27, 29, 31)) |>
  drop_na(E1F1r) |>
  pivot_longer(-SampleId,
               names_to = "Sample",
               values_to = "Value") |>
  pivot_wider(names_from = SampleId,
              values_from = Value) |>
  clean_names() |>
  mutate(across(-sample, as.double)) |>
  mutate(id2 = str_extract(sample, "^[WE]\\d[FN]")) |>
  group_by(id2) |>
  summarise(
    across(where(is.numeric), mean, na.rm = T
    
  ))
```

Setting units, in this code also main calculations of new variables is performed, and the numbers are later strip from their variables so they are compatibles with statistical analysis functions.

```{r}
area_m2 <- set_units(0.02, "m^2")
hyp_data <- all_data |>
  select(c(1:8, 16, 19:23, 25:27)) |>
  mutate(weight_g = set_units(weight, "g"),
         thickness_um = set_units(thickness,"um"),
         force_N = set_units(`F_[N]`, "N"),
         # Total_yield = set_units(Total_yield*100, "%"),
         # Total_screened_yield = set_units(Total_screened_yield*100, "%"),
         ModElas_MPa = set_units(`ModElas_[MPa]`, "MPa"),
         tear_gf = set_units(tear_gf, "gf")
         ) |>
  mutate(bw_g_m2 = weight_g * 0.93 / area_m2,
         density_g_cm3 = set_units(bw_g_m2 / thickness_um, "g/cm^3"),
         T_index_kNm_kg = set_units(force_N/ set_units(15, "mm") / bw_g_m2,
                                    "kN*m/kg"),
         Code = factor(Code),
         tear_N = set_units(tear_gf, "N"),
         tear_index_m2N_kg = set_units (tear_N / bw_g_m2, "m^2*N/kg")
         ) |>
  drop_units() |>
  select(-c("F_[N]", "tFmax [s]", "ModElas_[MPa]",
            "weight", "thickness", "TEA_[J/m²]"))
  #|>
  # group_by(
  #   Code,
  #   biomass,
  #   fines,
  #   hs_treatment
  # ) #|>
  # summarise(
  #   T_index = mean(T_index, na.rm = TRUE),
  #   n_sheets = n(),
  #   .groups = "drop"
  # )
```

```{r}

library(ggplot2)

ggplot(hyp_data |> filter(hs_treatment == "additive"),
       aes(x = density_g_cm3,
           y = T_index_kNm_kg,
           color = biomass
           )) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE) +
  labs(
    x = "Density g/cm^3",
    y = "T Index",
    color = "Treatment"
  ) +
  theme_minimal()


```


Running for hypothesis test

First we test whether fines affect tensile or not

$H_o: \mu_{Fines} = \mu_{NoFines}$

$H_a: \mu_{Fines} \neq \mu_{NoFines}$

```{r}
mod_full <- lmer(
  T_index_kNm_kg ~ biomass * fines * hs_treatment +
    (1 | Code),
  data = hyp_data
)

anova(mod_full, type = 3)
```

Based on the p-values, we reject the null hypothesis, and have strong evidence to conclude that the probability of getting this value under the assumption that the null hypothesis is true, is very low. Therefore we conclude that Fines do have an effect on the tensile index of the handsheets.

Also, based on the p-value for the interaction between biomass:fines, we have strong evidence to conclude that there is an interaction between biomass and fines, so the effect of fines differs between eucalyptus and wheat straw.

Based on th p-value for the interaction between fines and the handsheet treatment, we fail to reject the null hypothesis, and conclude that the handsheet treatment does not change the effect of the fines (no direct correlation between the treatment and the fines type)

And finally based on the interaction term between biomass, fines, and treatment, we can conclude that the use of the cationic polymer, or the re circulation does not affect differently the wheat straw of the eucalyptus.

```{r}
hs_data <- hyp_data %>%
  mutate(
    Code = factor(Code),

    biomass = factor(
      biomass,
      levels = c("E", "W"),
      labels = c("Eucalyptus", "Wheat_straw")
    ),

    fines = factor(
      fines,
      levels = c("N", "Y"),
      labels = c("No_fines", "Fines")
    ),

    hs_treatment = factor(
      hs_treatment,
      levels = c(
        "control",
        "recirculation",
        "additive"
      ),
      labels = c(
        "Control",
        "Recirculation",
        "Polymer"
      )
    )
  )
```

```{r}
sample_size_table <- hs_data |>
  group_by(
    biomass,
    Code,
    fines,
    hs_treatment
  ) |>
  summarise(
    n_measured = sum(!is.na(T_index_kNm_kg)),
    n_missing = sum(is.na(T_index_kNm_kg)),
    .groups = "drop"
  )

sample_size_table
```

```{r}
cell_data <- hs_data %>%
  group_by(
    Code,
    biomass,
    fines,
    hs_treatment
  ) %>%
  summarise(
    n_handsheets = sum(!is.na(T_index_kNm_kg)),

    mean_T_index = if_else(
      n_handsheets > 0,
      mean(T_index_kNm_kg, na.rm = TRUE),
      NA_real_
    ),

    sd_T_index = if_else(
      n_handsheets > 1,
      sd(T_index_kNm_kg, na.rm = TRUE),
      NA_real_
    ),

    se_T_index = if_else(
      n_handsheets > 1,
      sd_T_index / sqrt(n_handsheets),
      NA_real_
    ),

    .groups = "drop"
  )
```


```{r}
library(ggplot2)

ggplot(cell_data,
       aes(x = biomass,
           y = mean_T_index,
           fill = hs_treatment)) +
  geom_violin() +
  labs(
    x = "Biomass",
    y = "T Index",
    fill = "Treatment"
  ) +
  theme_minimal()

```

Descriptive statistics

```{r}
results_summary <- cell_data %>%
  group_by(
    biomass,
    fines,
    hs_treatment
  ) %>%
  summarise(
    n_cooks = sum(!is.na(mean_T_index)),
    mean = mean(mean_T_index, na.rm = TRUE),
    sd_between_cooks = sd(mean_T_index, na.rm = TRUE),
    se_between_cooks = sd_between_cooks / sqrt(n_cooks),
    .groups = "drop"
  )

results_summary
```

Cook level responses
```{r}
ggplot(
  cell_data,
  aes(
    x = fines,
    y = mean_T_index,
    color = biomass,
    group = interaction(biomass, Code)
  )
) +
  geom_line(alpha = 0.60) +
  geom_point(size = 2.8) +
  facet_wrap(~ hs_treatment) +
  labs(
    x = "Fines condition",
    y = "Mean tensile index",
    color = "Biomass"
  ) +
  theme_bw()
```

An upward line from No_fines to Fines means fines increased tensile index.

A downward line means fines decreased tensile index.

Different slopes for Eucalyptus and wheat straw suggest a biomass × fines interaction.

Different slope patterns across the three panels suggest a three-way interaction.

```{r}
mod_cook <- lmer(
  mean_T_index ~ biomass * fines * hs_treatment +
    (1 | Code),
  data = cell_data,
  REML = TRUE,
  na.action = na.exclude
)
```

