---
title: "Data processing"
author: "Yefrid Cordoba"
format: html
---

First import the table with weights and thickness

```{r}
#| message: False
#| warning: False
library(tidyverse)
library(readxl)
library(stringr)
library(janitor)
```

Now we start with the data 

```{r}
fines_data <- read_xlsx("raw_data/Fines_study.xlsx", 
                        sheet = "Handsheet_properties")
```

Extract the pulping data, and just select the kappa number

Here i can select which columns are going to appear in my final table
```{r}
Pulping_data <- read_xlsx("raw_data/Fines_study.xlsx", 
                        sheet = "Pulping") #|>
  # select(Biomass, Code, Kappa, Total_screened_yield, )
```

Here i can select which columns are going to appear in my final table
```{r}
freeness_data <- read_xlsx("raw_data/Fines_study.xlsx", 
                        sheet = "Freeness") |>
  select(ID1, freeness)
```

```{r}
Pulping_data
```

Getting the data from the tensile tests

```{r}
tensile_data <- read_xlsx("raw_data/tensile_data.xlsx")
tensile_data
```


We need to create the variables for each of the columns
```{r}
fines_data |>
  mutate(biomass = as.factor(Biomass),
         Biomass = NULL,
         fines = factor(ifelse(
           str_detect(ID,"F"), "Y", "N")),
         hs_treatment = factor(
           case_when(
            str_detect(ID, "A") ~ "additive",
            str_detect(ID, "5") ~ "recirculation",
            str_detect(ID, "[1-4]") ~ "control"),
           levels = c("control", "recirculation", "additive")),
         Code = str_extract(ID, "^[WE]\\d"),
         ID1 = str_extract(ID, "^[WE]\\d.")
         ) |>
  inner_join(Pulping_data, by = join_by(Code)) |>
  left_join(freeness_data, by = join_by(ID1)) |>
  left_join(tensile_data, by = join_by(ID)) |>
  select(-Biomass, -ID1) |>
  relocate(Code, biomass, fines, hs_treatment, .after = ID) |>
  mutate(
    id2 = str_extract(str_remove(ID, "A"), "^.{3}")
  )
  
```

Now we get a code to extract the data from the FQA values

```{r}
fqa_data <- read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/E1F1.xlsm",
                       sheet = "FQAResult",
                       range = "A8:B56")|>
  mutate((read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/E1F2.xlsm",
                       sheet = "FQAResult",
                       range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/E1N1.xlsm",
                       sheet = "FQAResult",
                       range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/E1N2.xlsm",
                       sheet = "FQAResult",
                       range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/E2F1.xlsm",
                       sheet = "FQAResult",
                       range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/E2F2.xlsm",
                       sheet = "FQAResult",
                       range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/E2N1.xlsm",
                       sheet = "FQAResult",
                       range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/E2N2.xlsm",
                       sheet = "FQAResult",
                       range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/E3F1.xlsm",
                       sheet = "FQAResult",
                       range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/E3F2.xlsm",
                       sheet = "FQAResult",
         range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/E3N2.xlsm",
                       sheet = "FQAResult",
         range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/E3N3.xlsm",
                       sheet = "FQAResult",
         range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/W1F1.xlsm",
                       sheet = "FQAResult",
         range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/W1F2.xlsm",
                       sheet = "FQAResult",
         range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/W1N1.xlsm",
                       sheet = "FQAResult",
         range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/W1N2.xlsm",
                       sheet = "FQAResult",
         range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/W2N1.xlsm",
                       sheet = "FQAResult",
         range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/W2N2.xlsm",
                       sheet = "FQAResult",
         range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/W3F1.xlsm",
                       sheet = "FQAResult",
         range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/W3F2.xlsm",
                       sheet = "FQAResult",
         range = "B8:B56")),
         (read_excel("C:/Users/yacordob/OneDrive - North Carolina State University/Fines Study Documents/FQA Documents/W3N1.xlsm",
                       sheet = "FQAResult",
         range = "B8:B56"))
)|>
  drop_na(SampleId)

```

Now that all the data is in that table, we need to fix the data

```{r}
fqa_data <- fqa_data|>
  slice(-c(1:13, 19, 22, 23, 27, 29, 31)) |>
  drop_na(E1F1r) |>
  pivot_longer(-SampleId,
               names_to = "Sample",
               values_to = "Value") |>
  pivot_wider(names_from = SampleId,
              values_from = Value) |>
  clean_names() |>
  mutate(across(-sample, as.double)) |>
  mutate(id2 = str_extract(sample, "^[WE]\\d[FN]")) |>
  group_by(id2) |>
  summarise(
    across(where(is.numeric), mean, na.rm = T
    
  ))
```

