---
title: "Data processing"
author: "Yefrid Cordoba"
format: html
---

## Libraries

```{r}
#| message: False
#| warning: False
library(tidyverse)
library(readxl)
library(stringr)
library(janitor)
library(units)
library(emmeans)
library(lmerTest)
library(lme4)
library(performance)
library(see)
library(DHARMa)
library(pbkrtest)
library(ggpmisc)
```

## Importing data


### 1. Handsheet data

Now we start with the data 

This first code chunk will import the data of the handsheets that was measured in the lab like weight, caliper, and tear.

```{r}
fines_data <- read_xlsx("raw_data/Fines_study.xlsx", 
                        sheet = "Handsheet_properties")
```

### 2. Pulping data

Import the pulping data like temperature, time, chemical charge, pulping, and screening yield, and kappa.

Here i can select which columns are going to appear in my final table
```{r}
Pulping_data <- read_xlsx("raw_data/Fines_study.xlsx", 
                        sheet = "Pulping") #|>
  # select(Biomass, Code, Kappa, Total_screened_yield, )
```

### 3. Freeness data

This code is importing the freeness data and selecting just the final values and an ID to later related to other tables.
```{r}
freeness_data <- read_xlsx("raw_data/Fines_study.xlsx", 
                        sheet = "Freeness") |>
  select(ID1, freeness)
```

### 4. Tensile data

In this code chunk we are going to import the data from the tensile test.

Getting the data from the tensile tests

```{r}
tensile_data <- read_xlsx("raw_data/tensile_data.xlsx")

```

### 5. FQA data

We are going to call the file where previously imported data is stored, this is done in this way, since the amount of excel files that needs to be read each time is big, and create computation constrains. In case the data does not exist, execute the file `fqa_data_extraction.R`

After the `.rds` file has been created, we are going to bring it to our environment as a new object.

```{r}
fqa_data <- readRDS("raw_data/fqa_data.rds")

```

## Wrangling data

All the previously imported have different variables, columns, spaces, missing values, etc, which need to be fixed below.

### 1. Master table

Creating a master table with all the data that is going to be used later
We need to create the variables for each of the columns
```{r}
all_data <- fines_data |>
  mutate(biomass = as.factor(Biomass),
         Biomass = NULL,
         fines = factor(ifelse(
           str_detect(ID,"F"), "Y", "N")),
         hs_treatment = factor(
           case_when(
            str_detect(ID, "A") ~ "additive",
            str_detect(ID, "5") ~ "recirculation",
            str_detect(ID, "[1-4]") ~ "control"),
           levels = c("control", "recirculation", "additive")),
         Code = str_extract(ID, "^[WE]\\d"),
         ID1 = str_extract(ID, "^[WE]\\d.")
         ) |>
  inner_join(Pulping_data, by = join_by(Code)) |>
  left_join(freeness_data, by = join_by(ID1)) |>
  left_join(tensile_data, by = join_by(ID)) |>
  select(-Biomass, -ID1) |>
  relocate(Code, biomass, fines, hs_treatment, .after = ID) |>
  mutate(
    id2 = str_extract(str_remove(ID, "A"), "^.{3}")
  )
  
```

###2. FQA wrangling

Now that all the data is in that table, we need to fix the data

```{r}
fqa2_data <- fqa_data|>
  slice(-c(1:13, 19, 22, 23, 27, 29, 31)) |>
  drop_na(E1F1r) |>
  pivot_longer(-SampleId,
               names_to = "Sample",
               values_to = "Value") |>
  pivot_wider(names_from = SampleId,
              values_from = Value) |>
  clean_names() |>
  mutate(across(-sample, as.double)) |>
  mutate(id2 = str_extract(sample, "^[WE]\\d[FN]")) |>
  group_by(id2) |>
  summarise(
    across(where(is.numeric), mean, na.rm = T
    
  ))
```



### 3. Hypothesis test data

Here we are setting units, in this code also main calculations of new variables is performed, and the numbers are later strip from their units so they are compatibles with statistical analysis functions.

```{r}
area_m2 <- set_units(0.02, "m^2")
hyp_data <- all_data |>
  select(c(1:8, 16, 19:23, 25:27)) |>
  mutate(weight_g = set_units(weight, "g"),
         thickness_um = set_units(thickness,"um"),
         force_N = set_units(`F_[N]`, "N"),
         # Total_yield = set_units(Total_yield*100, "%"),
         # Total_screened_yield = set_units(Total_screened_yield*100, "%"),
         ModElas_MPa = set_units(`ModElas_[MPa]`, "MPa"),
         tear_gf = set_units(tear_gf, "gf")
         ) |>
  mutate(bw_g_m2 = weight_g * 0.93 / area_m2,
         density_g_cm3 = set_units(bw_g_m2 / thickness_um, "g/cm^3"),
         T_index_kNm_kg = set_units(force_N/ set_units(15, "mm") / bw_g_m2,
                                    "kN*m/kg"),
         Code = factor(Code),
         tear_N = set_units(tear_gf, "N"),
         tear_index_m2N_kg = set_units (tear_N / bw_g_m2, "m^2*N/kg")
         ) |>
  drop_units() |>
  select(-c("F_[N]", "tFmax [s]", "ModElas_[MPa]",
            "weight", "thickness", "TEA_[J/m²]"))
  #|>
  # group_by(
  #   Code,
  #   biomass,
  #   fines,
  #   hs_treatment
  # ) #|>
  # summarise(
  #   T_index = mean(T_index, na.rm = TRUE),
  #   n_sheets = n(),
  #   .groups = "drop"
  # )
```

## Exploratory Data Analysis

In this section we are going to perform basic exploratory data analysis before starting with the hypothesis test.

We are going to check distributions, with boxplots, bars charts, and check normality.

### 1. Checking normality



```{r}

library(ggplot2)

ggplot(hyp_data, # |> filter(hs_treatment == "control"),
       aes(x = density_g_cm3,
           y = tear_index_m2N_kg,
           color = hs_treatment
           )) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE) +
  stat_poly_eq(
aes(
label = paste(..rr.label..)
),
formula = y ~ x,
parse = TRUE
) +
  labs(
    x = "Density g/cm^3",
    y = "T Index",
    color = "Treatment"
  ) +
  theme_minimal()


```


Running for hypothesis test

First we test whether fines affect tensile or not

$H_o: \mu_{Fines} = \mu_{NoFines}$

$H_a: \mu_{Fines} \neq \mu_{NoFines}$

```{r}
mod_full <- lmer(
  T_index_kNm_kg ~ biomass * fines * hs_treatment +
    (1 | Code),
  data = hyp_data
)

anova(mod_full, type = 3)
```

Based on the p-values, we reject the null hypothesis, and have strong evidence to conclude that the probability of getting this value under the assumption that the null hypothesis is true, is very low. Therefore we conclude that Fines do have an effect on the tensile index of the handsheets.

Also, based on the p-value for the interaction between biomass:fines, we have strong evidence to conclude that there is an interaction between biomass and fines, so the effect of fines differs between eucalyptus and wheat straw.

Based on th p-value for the interaction between fines and the handsheet treatment, we fail to reject the null hypothesis, and conclude that the handsheet treatment does not change the effect of the fines (no direct correlation between the treatment and the fines type)

And finally based on the interaction term between biomass, fines, and treatment, we can conclude that the use of the cationic polymer, or the re circulation does not affect differently the wheat straw of the eucalyptus.

```{r}
hs_data <- hyp_data %>%
  mutate(
    Code = factor(Code),

    biomass = factor(
      biomass,
      levels = c("E", "W"),
      labels = c("Eucalyptus", "Wheat_straw")
    ),

    fines = factor(
      fines,
      levels = c("N", "Y"),
      labels = c("No_fines", "Fines")
    ),

    hs_treatment = factor(
      hs_treatment,
      levels = c(
        "control",
        "recirculation",
        "additive"
      ),
      labels = c(
        "Control",
        "Recirculation",
        "Polymer"
      )
    )
  )
```

```{r}
sample_size_table <- hs_data |>
  group_by(
    biomass,
    Code,
    fines,
    hs_treatment
  ) |>
  summarise(
    n_measured = sum(!is.na(T_index_kNm_kg)),
    n_missing = sum(is.na(T_index_kNm_kg)),
    .groups = "drop"
  )

sample_size_table
```

```{r}
cell_data <- hs_data %>%
  group_by(
    Code,
    biomass,
    fines,
    hs_treatment
  ) %>%
  summarise(
    n_handsheets = sum(!is.na(T_index_kNm_kg)),

    mean_T_index = if_else(
      n_handsheets > 0,
      mean(T_index_kNm_kg, na.rm = TRUE),
      NA_real_
    ),

    sd_T_index = if_else(
      n_handsheets > 1,
      sd(T_index_kNm_kg, na.rm = TRUE),
      NA_real_
    ),

    se_T_index = if_else(
      n_handsheets > 1,
      sd_T_index / sqrt(n_handsheets),
      NA_real_
    ),

    .groups = "drop"
  )
```


```{r}
library(ggplot2)

ggplot(cell_data,
       aes(x = biomass,
           y = mean_T_index,
           fill = hs_treatment)) +
  geom_violin() +
  labs(
    x = "Biomass",
    y = "T Index",
    fill = "Treatment"
  ) +
  theme_minimal()

```

Descriptive statistics

```{r}
results_summary <- cell_data %>%
  group_by(
    biomass,
    fines,
    hs_treatment
  ) %>%
  summarise(
    n_cooks = sum(!is.na(mean_T_index)),
    mean = mean(mean_T_index, na.rm = TRUE),
    sd_between_cooks = sd(mean_T_index, na.rm = TRUE),
    se_between_cooks = sd_between_cooks / sqrt(n_cooks),
    .groups = "drop"
  )

results_summary
```

Cook level responses
```{r}
ggplot(
  cell_data,
  aes(
    x = fines,
    y = mean_T_index,
    color = biomass,
    group = interaction(biomass, Code)
  )
) +
  geom_line(alpha = 0.60) +
  geom_point(size = 2.8) +
  facet_wrap(~ hs_treatment) +
  labs(
    x = "Fines condition",
    y = "Mean tensile index",
    color = "Biomass"
  ) +
  theme_bw()
```

An upward line from No_fines to Fines means fines increased tensile index.

A downward line means fines decreased tensile index.

Different slopes for Eucalyptus and wheat straw suggest a biomass × fines interaction.

Different slope patterns across the three panels suggest a three-way interaction.

```{r}
mod_cook <- lmer(
  mean_T_index ~ biomass * fines * hs_treatment +
    (1 | Code),
  data = cell_data,
  REML = TRUE,
  na.action = na.exclude
)
```

